### 2021-08-06 based on ggpubr/examples of ggmatplot.Rmd
latest ggmatplot functions depending on package ggpubr

---
title: "Examples of ggmatplot"
author: "Xuan Liang"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: journal
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

```{r setup data, message=FALSE}
library(tidyverse)
library(ggpubr)

source("ggmatlist.R") # The function is to convert a list to a matrix
source("ggmatclean.R") # The function is to transfrom a wide format data to a long format data
source("ggmatplot.R") # Plot function without ... option as an argument
source("getoption.R") # Used in ggmatplot1() function to extract what arguments can be allowed in a function of ggpubr, for example, I only tried it for plot.type "histogram". 
source("ggmatplot1.R") # Plot1 function with ... option as an argument. It seems here are some issues ralated with ... option.

# Default plot type: Scatterplot
x <- rnorm(100,sd=2)
y <- x*0.5+rnorm(100,sd=1)
fit.y <- fitted(lm(y~x))
p <- ggmatplot(x,cbind(y,fit.y)) 
p
class(p)

# Plot.type = "line" or "both"
X <- 1:10
Y <- X^2
Z <- X^2-2*X

ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),plot.type = "line")
ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),plot.type = "both")

# Plot.type = "density" or "histogram" or "boxplot"
iris_sub <- subset(iris, Species == "setosa")

ggmatplot(iris_sub[1:4], plot.type = "density")

ggmatplot(iris_sub[1:4], plot.type = "histogram")

ggmatplot(iris_sub[1:4], plot.type = "boxplot")

ggmatplot(iris_sub[1:4], plot.type = "dotplot")

ggmatplot(iris_sub[1:2], plot.type = "errorplot")

ggmatplot(iris_sub[1:4], plot.type = "violin")

ggmatplot(iris_sub[1:4], plot.type = "ecdf", xlab="Length", ylab="ecdf")

# Use a list as input
x <- list(low = rnorm(30, -5, 2),
          medium = rnorm(20),
          high = rnorm(40, 3, 0.5))

ggmatplot(x, plot.type = "both", legend_title = "Level")

ggmatplot(x, plot.type = "density", legend_title = "Level")

# Test if the ... option in ggmatplot1 works well
# Correct additional argument
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, rug=TRUE)

# Wrong additional argument but no errors or warnings show up
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, rugg=TRUE)
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, bins=20)
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, binwdith=20)

ggmatplot(x, plot.type = "histogram", legend_title = "Level")
ggmatplot1(x, plot.type = "histogram",legend_title = "Level", bins=20)
ggmatplot1(x, plot.type = "histogram",legend_title = "Level", bins=20, rug=TRUE, add="mean",alpha=0.3)

ggmatplot(x, plot.type = "boxplot", legend_title = "Level")

# Data
stocks <- data.frame(
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4)
)
stocklong <- stocks %>% gather(stock, price)

# some nice features in ggpubr
p1 <- ggboxplot(stocklong, x = "stock", y = "price",add = "jitter",color="stock", shape = "stock")
p1

my_comparisons <- list( c("X", "Y"), c("Y", "Z"), c("X", "Z") )
p1 + stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 12)

# test the features in ggmatplot
ggmatplot(stocks, plot.type = "boxplot", legend_title = "Level")+ 
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 12)

```