### 2022-05-05 Paper/paper.Rmd
paper first draft done

---
title: '`ggmatplot`: An R package for data visualization on wide-format data'
tags:
- R
- grammar of graphics
- ggplot2
- data visualisation
authors:
  - name: Xuan Liang
    orcid: 0000-0002-0792-3751
    affiliation: 1
  - name: Francis K. C. Hui
    orcid: 0000-0003-0765-3533
    affiliation: 1
  - name: Dilinie Seimon
    orcid: 0000-0003-4431-6077
    affiliation: 2
  - name: Emi Tanaka
    orcid: 0000-0002-1455-259X
    affiliation: 2
affiliations:
  - name: Research School of Finance, Actuarial Studies and Statistics, The Australian National University
    index: 1
  - name: Department of Econometrics and Business Statistics, Monash University
    index: 2
citation_author: Liang et. al.
date: "`r Sys.Date()`"
year: 2022
bibliography: paper.bib
output: 
  rticles::joss_article:
    keep_tex: true
csl: apa.csl
journal: JOSS
nocite: '@mosaicData'
---

```{r setup, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# data

wide_df <- tribble(~restaurant, ~food_rating, ~service_rating, ~ambience_rating, ~overall_rating,
        "R1", 4.3, 3.4, 4.3, 4.9,
        "R2", 4.3, 5.0, 4.5, 4.4,
        "R3", 3.2, 4.4, 5.0, 3.0, 
        "R4", 2.3, 4.6, 4.4, 3.8, 
        "R5", 3.9, 4.8, 4.2, 3.3)

long_df <- pivot_longer(wide_df, contains("rating"), 
                        names_to = "rating_type",
                        names_pattern = "(.+)_rating",
                        values_to = "rating")

# abund_df <- tribble(~Site, ~`Soil dry mass`, ~`Moss`, ~Alopcune, ~Arctlute, ~Pardpull, ~Trocterr, ~Zoraspin,
#                     1, 2.3321, 3.0445, 10, 0, 45, 57, 4,
#                     2, 3.0493, 1.0986, 2, 0, 37, 65, 9,
#                     3, 2.5572, 2.3979, 20, 0, 45, 66, 1)
# 
table_styling <- function(df, ...) {
  kable(df, ..., booktabs = TRUE, linesep = "") %>%
    kable_classic()
}
# 
# data("spider", package = "mvabund")

data(SnowGR, package = "mosaicData")

```

# Summary

The layered grammar of graphics [@Wickham2010-kt], implemented as the `ggplot2` package [@Wickham2016] in the statistical language R [@rstats], is a powerful and popular tool to create versatile statistical graphics. This graphical system, however, requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g. x-coordinate, y-coordinate, color, size), which create friction in constructing plots with an aesthetic element that span multiple columns in the original data by requiring users to re-organise the data.

The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrices or data with the result returned as a `ggplot` object. The package is inspired by the function `matplot()` in the core R `graphics` system, thus `ggmatplot` can be considered as a `ggplot` version of `matplot` with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions.

# Statement of need

Input data to construct plots with `ggplot2` require data to be organised in a manner that maps data columns to aesthetic elements. This generally works well where data is tidied in a rectangular form, referred to as "tidy data" [@Wickham2014-gy], where each row represents an observational unit, each column represents a variable, and each cell represents a value. In some cases, what constitutes a variable (or observational unit), hence a column (or row), in a tidy data can be dependent upon interpretation or downstream interest (e.g. Tables \ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data), but a clear violation of tidy data principles is when the column names contain data values, e.g. Table \ref{tab:tab3} contain the name of the species across a number of column names.

```{r tab1, echo = FALSE}
table_styling(wide_df, col.names = c("Restaurant", "Food",
                                     "Service", "Ambience", "Overall"),
              caption = 'Restaurant rating data in "tidy" form. The first column shows the restaurant ID, and the next four columns show the average ratings (out of 5) for food, service, ambience and overall, respectively.') %>% 
  add_header_above(c(" ", "Average rating" = 4))
```

```{r tab2, echo = FALSE}
table_styling(long_df, col.names = c("Restauant", "Rating type", "Average rating"),
              caption = 'Another form for the restaurant rating data in Table \\ref{tab:tab1}. In Wickham (2014), this format is called the "molten" data.')
```

```{r tab3, echo = FALSE}
head(SnowGR[, 1:11]) %>% 
  mutate(Jan = as.character(Jan)) %>% 
  replace_na(list(Jan = ".")) %>% 
  table_styling(caption = 'The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the `mosaicData` R pacakge (Prium, Kaplan \\& Horton, 2021).',
                align = "r")
```

The organisation of the data is largely dependent on the downstream analysis and there is no one correct way to do this. Some forms of multivariate data, e.g. Table \ref{tab:tab3}, are prevalent in the field because it aligns as an input data for a modelling software and/or the format is more convenient for input or view of the data in spreadsheet format. However, this format is not consistent with the required format for `ggplot2`, and consequently, plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly visualise these types of data. The `ggmatplot` R-package provides a solution to this common friction in producing plots with `ggplot2`.

# Examples

In this section we demonstrate the use of the `ggmatplot` package and contrast the specification with `ggplot2` after data wrangling using `dplyr` and `tidyr` [@Wickham2019] using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3}, which are stored in the objects `wide_df` and `SnowGR`, respectively.

## Example 1

The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of rating, contained in columns 2 to 4 of `wide_df`, against the overall restaurant rating in column 5 of `wide_df` as shown in Figure 1.

```{r matplot2, fig.height = 2, fig.cap = "Line plot of the food, service and ambience rating against the overall restaurant rating."}
library(ggmatplot)
ggmatplot(x = wide_df[, 2:4], y = wide_df[, 5], plot_type = "both",
          xlab = "Rating",  ylab = "Overall rating", legend_title = "Type") 
```

In contrast to the above, the data must be wrangled to a long form first before plotting in the code below to obtain the same figure as Figure 1. This adds a small, but noticable, friction to the workflow for the user that is looking to quickly explore their data.

```{r ggplot2, fig.height = 2, eval = FALSE}
library(ggplot2)
library(tidyr) # or library(tidyverse)
wide_df %>% 
  select(contains("rating")) %>% 
  pivot_longer(-overall_rating, 
               names_to = "rating_type",
               values_to = "rating") %>% 
  ggplot(aes(rating, overall_rating, color = rating_type)) + 
  geom_point(aes(shape = rating_type)) +
  geom_line(aes(group = rating_type, linetype = rating_type)) +
  labs(x = "Restaurant", y = "Rating", 
       color = "Type", linetype = "Type", shape = "Type")
```

## Example 2

The example code draws the boxplot of each column of amount of snowfall across months in the `SnowGR` data as shown in Figure 2. As the resulting object is a `ggplot` object, the user can leverage the `ggplot` functions to modify the output (e.g. removal of the legend).

```{r matplot3, fig.height = 2, fig.cap="The distribution of the amount of snowfall at Grand Rapids, Michigan, across months from 1893-2011."}
library(ggmatplot)
ggmatplot(x = SnowGR[, 3:14], plot_type = "boxplot",
          xlab = "Month",  ylab = "Snowfall") +
  guides(color = "none", fill = "none")
```

The equivalent code for the above to produce Figure 2 without using `ggmatplot` is given below. There is again a slight friction in putting the data in the right format before plotting. The original wide data format like those shown in Table \ref{tab:tab3} is common so an analyst who has to repeat these tasks can benefit from a quick approach.

```{r ggplot3, fig.height = 2, eval = FALSE}
library(ggplot2)
library(tidyr) # or library(tidyverse)
SnowGR %>% 
  pivot_longer(Jul:Jun, 
               names_to = "Month",
               values_to = "Snowfall") %>% 
  ggplot(aes(Month, Snowfall)) + 
  geom_boxplot(aes(color = Month, fill = Month), alpha = 0.5) +
  guides(color = "none", fill = "none")
```

# Discussion

The `ggmatplot` R-package provides a solution to a common friction to quickly plotting multivariate data where the primary interest is mapping the column names as an aesthetic element. The solution provided however is a recipe-driven approach where the user can only produce plot types as many there are included in the `plot_type` option. Future development of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.

# Acknowledgements

FKCH was supported by ARC DECRA XXX.

# References