### 2022-05-05 Paper/paper.tex

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
% \renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/4.1/Resources/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\footnotesize{\sffamily Liang et.
al., (2022). \texttt{ggmatplot}: An R package for data visualization on
wide-format
data. \textit{Journal of Open Source Software}, (), . \href{https://doi.org/}{https://doi.org/}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={ggmatplot: An R package for data visualization on wide-format data},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{\texttt{ggmatplot}: An R package for data visualization on
wide-format data}

        \author[1]{Xuan Liang}
          \author[1]{Francis K. C. Hui}
          \author[2]{Dilinie Seimon}
          \author[2]{Emi Tanaka}

      \affil[1]{Research School of Finance, Actuarial Studies and
Statistics, The Australian National University}
      \affil[2]{Department of Econometrics and Business Statistics,
Monash University}
  \date{\vspace{-5ex}}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/}{\color{linky}{}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{}{\color{linky}{Review}} \ExternalLink
    \item \href{}{\color{linky}{Repository}} \ExternalLink
    \item \href{}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  {\bfseries Submitted:} \\
  {\bfseries Published:}

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC-BY}}).
}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

The layered grammar of graphics (Wickham, 2010), implemented as the
\texttt{ggplot2} package (Wickham, 2016) in the statistical language R
(R Core Team, 2021), is a powerful and popular tool to create versatile
statistical graphics. This graphical system, however, requires input
data to be organised in a manner that a data column is mapped to an
aesthetic element (e.g.~x-coordinate, y-coordinate, color, size), which
create friction in constructing plots with an aesthetic element that
span multiple columns in the original data by requiring users to
re-organise the data.

The \texttt{ggmatplot}, built upon \texttt{ggplot2}, is an R-package
that allows quick plotting across the columns of matrices or data with
the result returned as a \texttt{ggplot} object. The package is inspired
by the function \texttt{matplot()} in the core R \texttt{graphics}
system, thus \texttt{ggmatplot} can be considered as a \texttt{ggplot}
version of \texttt{matplot} with the benefits of customising the plots
as any other \texttt{ggplot} objects via \texttt{ggplot2} functions.

\hypertarget{statement-of-need}{%
\section{Statement of need}\label{statement-of-need}}

Input data to construct plots with \texttt{ggplot2} require data to be
organised in a manner that maps data columns to aesthetic elements. This
generally works well where data is tidied in a rectangular form,
referred to as ``tidy data'' (Wickham, 2014), where each row represents
an observational unit, each column represents a variable, and each cell
represents a value. In some cases, what constitutes a variable (or
observational unit), hence a column (or row), in a tidy data can be
dependent upon interpretation or downstream interest (e.g.~Tables
\ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data),
but a clear violation of tidy data principles is when the column names
contain data values, e.g.~Table \ref{tab:tab3} contain the name of the
species across a number of column names.

\begin{table}

\caption{\label{tab:tab1}Restaurant rating data in "tidy" form. The first column shows the restaurant ID, and the next four columns show the average ratings (out of 5) for food, service, ambience and overall, respectively.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Average rating} \\
\cmidrule(l{3pt}r{3pt}){2-5}
Restaurant & Food & Service & Ambience & Overall\\
\midrule
R1 & 4.3 & 3.4 & 4.3 & 4.9\\
R2 & 4.3 & 5.0 & 4.5 & 4.4\\
R3 & 3.2 & 4.4 & 5.0 & 3.0\\
R4 & 2.3 & 4.6 & 4.4 & 3.8\\
R5 & 3.9 & 4.8 & 4.2 & 3.3\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:tab2}Another form for the restaurant rating data in Table \ref{tab:tab1}. In Wickham (2014), this format is called the "molten" data.}
\centering
\begin{tabular}[t]{llr}
\toprule
Restauant & Rating type & Average rating\\
\midrule
R1 & food & 4.3\\
R1 & service & 3.4\\
R1 & ambience & 4.3\\
R1 & overall & 4.9\\
R2 & food & 4.3\\
R2 & service & 5.0\\
R2 & ambience & 4.5\\
R2 & overall & 4.4\\
R3 & food & 3.2\\
R3 & service & 4.4\\
R3 & ambience & 5.0\\
R3 & overall & 3.0\\
R4 & food & 2.3\\
R4 & service & 4.6\\
R4 & ambience & 4.4\\
R4 & overall & 3.8\\
R5 & food & 3.9\\
R5 & service & 4.8\\
R5 & ambience & 4.2\\
R5 & overall & 3.3\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:tab3}The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the `mosaicData` R pacakge (Prium, Kaplan \& Horton, 2021).}
\centering
\begin{tabular}[t]{rrrrrrrrrrr}
\toprule
SeasonStart & SeasonEnd & Jul & Aug & Sep & Oct & Nov & Dec & Jan & Feb & Mar\\
\midrule
1893 & 1894 & 0 & 0 & 0 & 0.0 & 8.0 & 24.9 & 12.5 & 6.8 & 4.8\\
1894 & 1895 & 0 & 0 & 0 & 0.0 & 7.5 & 5.3 & 21.5 & 8.0 & 22.5\\
1895 & 1896 & 0 & 0 & 0 & 0.4 & 23.2 & 15.0 & . & 8.5 & 2.0\\
1896 & 1897 & 0 & 0 & 0 & 0.2 & 8.0 & 8.0 & 4.9 & 11.2 & 12.0\\
1897 & 1898 & 0 & 0 & 0 & 0.0 & 1.4 & 8.0 & 15.5 & 29.5 & 0.0\\
1898 & 1899 & 0 & 0 & 0 & 0.0 & 18.5 & 18.0 & 20 & 3.4 & 16.0\\
\bottomrule
\end{tabular}
\end{table}

The organisation of the data is largely dependent on the downstream
analysis and there is no one correct way to do this. Some forms of
multivariate data, e.g.~Table \ref{tab:tab3}, are prevalent in the field
because it aligns as an input data for a modelling software and/or the
format is more convenient for input or view of the data in spreadsheet
format. However, this format is not consistent with the required format
for \texttt{ggplot2}, and consequently, plotting with \texttt{ggplot2}
interrupts the workflow of a user that is trying to quickly visualise
these types of data. The \texttt{ggmatplot} R-package provides a
solution to this common friction in producing plots with
\texttt{ggplot2}.

\hypertarget{examples}{%
\section{Examples}\label{examples}}

In this section we demonstrate the use of the \texttt{ggmatplot} package
and contrast the specification with \texttt{ggplot2} after data
wrangling using \texttt{dplyr} and \texttt{tidyr} (Wickham et al., 2019)
using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3},
which are stored in the objects \texttt{wide\_df} and \texttt{SnowGR},
respectively.

\hypertarget{example-1}{%
\subsection{Example 1}\label{example-1}}

The code below constructs a line plot (superimposed with a point) of the
various types (food, service and ambience) of rating, contained in
columns 2 to 4 of \texttt{wide\_df}, against the overall restaurant
rating in column 5 of \texttt{wide\_df} as shown in Figure 1.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggmatplot)}
\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\AttributeTok{y =}\NormalTok{ wide\_df[, }\DecValTok{5}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{,}
          \AttributeTok{xlab =} \StringTok{"Rating"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Overall rating"}\NormalTok{, }\AttributeTok{legend\_title =} \StringTok{"Type"}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/matplot2-1.pdf}
\caption{Line plot of the food, service and ambience rating against the
overall restaurant rating.}
\end{figure}

In contrast to the above, the data must be wrangled to a long form first
before plotting in the code below to obtain the same figure as Figure 1.
This adds a small, but noticable, friction to the workflow for the user
that is looking to quickly explore their data.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# or library(tidyverse)}
\NormalTok{wide\_df }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(}\FunctionTok{contains}\NormalTok{(}\StringTok{"rating"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{overall\_rating, }
               \AttributeTok{names\_to =} \StringTok{"rating\_type"}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{"rating"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(rating, overall\_rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{shape =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{group =}\NormalTok{ rating\_type, }\AttributeTok{linetype =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Restaurant"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Rating"}\NormalTok{, }
       \AttributeTok{color =} \StringTok{"Type"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"Type"}\NormalTok{, }\AttributeTok{shape =} \StringTok{"Type"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{example-2}{%
\subsection{Example 2}\label{example-2}}

The example code draws the boxplot of each column of amount of snowfall
across months in the \texttt{SnowGR} data as shown in Figure 2. As the
resulting object is a \texttt{ggplot} object, the user can leverage the
\texttt{ggplot} functions to modify the output (e.g.~removal of the
legend).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggmatplot)}
\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ SnowGR[, }\DecValTok{3}\SpecialCharTok{:}\DecValTok{14}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"boxplot"}\NormalTok{,}
          \AttributeTok{xlab =} \StringTok{"Month"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \StringTok{"none"}\NormalTok{, }\AttributeTok{fill =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/matplot3-1.pdf}
\caption{The distribution of the amount of snowfall at Grand Rapids,
Michigan, across months from 1893-2011.}
\end{figure}

The equivalent code for the above to produce Figure 2 without using
\texttt{ggmatplot} is given below. There is again a slight friction in
putting the data in the right format before plotting. The original wide
data format like those shown in Table \ref{tab:tab3} is common so an
analyst who has to repeat these tasks can benefit from a quick approach.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# or library(tidyverse)}
\NormalTok{SnowGR }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(Jul}\SpecialCharTok{:}\NormalTok{Jun, }
               \AttributeTok{names\_to =} \StringTok{"Month"}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Month, Snowfall)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_boxplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{color =}\NormalTok{ Month, }\AttributeTok{fill =}\NormalTok{ Month), }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \StringTok{"none"}\NormalTok{, }\AttributeTok{fill =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The \texttt{ggmatplot} R-package provides a solution to a common
friction to quickly plotting multivariate data where the primary
interest is mapping the column names as an aesthetic element. The
solution provided however is a recipe-driven approach where the user can
only produce plot types as many there are included in the
\texttt{plot\_type} option. Future development of the package could
benefit from using a grammar approach, like in Wilkinson (2005) and
Wickham (2010), where plot types can be extensible.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

FKCH was supported by ARC DECRA XXX.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-mosaicData}{}}%
Pruim, R., Kaplan, D., \& Horton, N. (2021). \emph{mosaicData: Project
MOSAIC data sets}. Retrieved from
\url{https://CRAN.R-project.org/package=mosaicData}

\leavevmode\vadjust pre{\hypertarget{ref-rstats}{}}%
R Core Team. (2021). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2010-kt}{}}%
Wickham, H. (2010). A layered grammar of graphics. \emph{Journal of
computational and graphical statistics: a joint publication of American
Statistical Association, Institute of Mathematical Statistics, Interface
Foundation of North America}.

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2014-gy}{}}%
Wickham, H. (2014). Tidy data. \emph{Journal of Statistical Software},
\emph{59}(10), 1--23.
doi:\href{https://doi.org/10.18637/jss.v059.i10}{10.18637/jss.v059.i10}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from
\url{https://ggplot2.tidyverse.org}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2019}{}}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., et al. (2019). Welcome to the {tidyverse}.
\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}

\leavevmode\vadjust pre{\hypertarget{ref-Wilkinson2005-oz}{}}%
Wilkinson, L. (2005). \emph{The grammar of graphics}. Springer.

\end{CSLReferences}

\end{document}